---
title: "code chunk demo"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


#Setup chunk

'''{r load_penguin_data, echo=TRUE, eval=TRUE}
require(palmerpenguins)
head(penguins)
'''

#my chunk
'''{r simple pairplot}
plot(penguins[, 3:5])
'''

#echo = FALSE will not show the code and wont be evaluated
#for fig.asp, 1 is a square plot and 1/1.5 is a rectangle
'''{r simple pairplot, fig.width=4, fig.asp = 1}
plot(penguins[, 3:5])
'''

'''{r simple pairplot, fig.width=4, fig.asp = 1/1.5}
plot(penguins[, 3:5])
'''

'''{r simple pairplot, eval = FALSE}
plot(penguins{, 3:5}}
'''


#plot will not appear in redered document


''''''{r simple pairplot, echo = FALSE}
plot(penguins[, 3:5})]
'''


# Tabset Example {.tabset}

## Tab 1
Tab 1 contents

## Tab 2
Tab 2 contents

## Tab 3
Tab 3 contents

# Next section
Non-tabbed material

#for a different looking tab

# Tabset Example {.tabset . tabset-pills}

###################

# Different Penguin Plots {.tabset . tabset-pills}

The following tabs contain different plot types


## Histogram

```{r flipper length histogram, fig. asp = 1/1.5}
hist(penguins$flipper_length_mm)
```



## Another Histogram (Bill Length)

```{r bill length histogram, echo =FALSE}
hist(b)
```














